Untwisting the Tetrahelix (v0.4)

### 2017-02-11 doc/TetrahelixUntwist.tex
some changes to scene and clarification

\begin{abstract}
  The Boerdijk--Coxeter helix (BC helix, or tetrahelix) is a face-to-face stack of regular tetrahedra whose outer vertices lie on helices.
  By allowing changes to edge lengths so that the tetrahedra are not regular, we can define a continuum of tetrahelices isomorphic to the
  BC helix of designable pitch or curvature, effectively ``untwisting'' the BC helix. We show that in the sense of minimal maximum
  difference in edge length, every tetrahelix has vertices equally spaced along the axis and at most three distinct edge lengths.
  In the case of a full untwisting to zero curvature, we describe a novel object, the
  \emph{equitetrabeam}.
  It has 3-fold symmetry about  axis, and chirality.  This tetrabeam and controllably twisted tetrahelices are
  interesting for structural engineering because they are inherently rigid space frames and trusses.
  A formula for the cartesian coordinates of the individual nodes of outer helices of the BC helix is given and
  unified with the formula for the equitetrabeam creating a class of edge-length optimal tetrahelices generated by single parameter.
  A further generalization is provided.
  Utility and use for truss/space frame design and robotics are discussed.
\end{abstract}

@@

\begin{theorem}
  \label{tetrahelixoptimality}
  Any one tetrahedron in a minimax-optimal tetrahelix $ABCD$ of edge lengths $AD = 1, AB = BC = CD = u , AC = BD = v$, where $1 \leq u \leq v$.
  Any one tetrahedron in a tetrahelix has $1$ rail edge, $2$ mid-length edges connected to the rail and $2$ long edges connected to the rail.
  The edge opposite of the rail edge is a mid-length edge.
\end{theorem}

@@

Note that based on Theorems \ref{tetrahelixoptimality} and \ref{eventhirds}, we are justified in classifying edge lengths as \emph{rail},\emph{mid-length}, or
\emph{long}. The mid-length edges are the edges between closest on the $z$-axis, and the long edges are those that hop over a vertex.

@@

We seek to unify this with degenerate helix formula for the equitetrabeam:
\[
H_{etb}(n,c) =
\left [
  \begin{tabular}{c}
   $ r_{etb}  \cos( 0 \cdot \kappa  + c 2 \pi /3) $\\
   $ r_{etb}  \sin( 0 \cdot \kappa  + c 2 \pi /3) $\\
   $ \kappa d_{etb} $
  \end{tabular}
\right ],
\text{where:}
  \begin{tabular}{c}
 $\kappa = n + c/3$ \\
    $\rho_{bc} = (3 \theta - 2 \pi)$ \\
   $ \theta = \arccos(-2/3) $
  \end{tabular}      
\]
where $ r_{etb} = 1/\sqrt(3)$, $d_{etb} = 1/3$,

\begin{align*}
r_{\lambda}  &=  \lambda (3 \sqrt(3) / 10 - 1/\sqrt(3)) +  1/\sqrt(3) \\
d_{\lambda} &=   \lambda (3 / \sqrt(10) - 1)+ 1 \\
\phi_{\lambda} &=  \lambda \rho_{bc}  + 0\\
\end{align*}
to create a formula that generates a continuum of tetrahedral structures:

\[
H_{continuum}(n,c,\lambda) =
\left [
  \begin{tabular}{c}
   $ r_{\lambda} \cos(\phi_{\lambda} \kappa + c 2 \pi /3) $\\
   $ r_{\lambda}  \sin(\phi_{\lambda} \kappa + c 2 \pi /3) $\\
   $ d_{\lambda} \kappa $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
    $\kappa = n + c/3$ \\
    $ \phi_{\lambda} =  \lambda \rho_{bc}  + 0 $\\
    $\rho_{bc} = (3 \theta - 2 \pi)$ \\
   $ \theta = \arccos(-2/3) $
  \end{tabular}      
\]
A value of $\lambda = 0$ generates the equitetrabeam, and $\lambda = 1$ generates the Boerdijk--Coxeter helix, and every
value $\lambda \in [0,1]$ generates an attractive structure which if physically realized is an inherently rigid structure
with member lengths of no greater disparity than $21\%$.
In the sense of structural engineering, it would be a relatively strong space frame.

Furthermore, this formula allows one to design the pitch (in edge length units) as a function of $\lambda$ of the helix (along one rail)
where $\lambda \neq 0$:
\[
p(\lambda) = 2 \pi  \cdot \frac{((3/\sqrt(10) -1) \lambda +1)}{ \lambda  \rho_{bc} }
\]

@@

Checking the important special case of the BC helix, we find that this equation
indeed holds true (treathing $d$ in this equation as $3 h_{bc}$ as defined by
Gray and Coxeter, where they are using it for the axial height from one node to
the next of a different color, but we use it to mean distance for the same color.

@@

By Theoerm \ref{eventhirds}, we can compute the (at most) three edge-lengths of an optimal
tetrahelix by (where $dist$ is the cartesian distance function):
\begin{align*}
  \text{rail} &= dist(H_{general}(n,c,\rho,h_{\rho}),H_{general}(n+1,c,\rho,d_{\rho})) = 1 \\
  \text{mid-length} &= dist(H_{general}(n,c,\rho,h_{\rho}),H_{general}(n,c+1,\rho,d_{\rho}))  \\
  \text{long} &= dist(H_{general}(n,c,\rho,h_{\rho}),H_{general}(n,c+2,\rho,d_{\rho}))  \\  
\end{align*}
Where are invarinat for all $n$ and $c$.

@@

When $\rho = 0$, our formulae for distances are particular simple. Let $a$ be the length of the equilateral triangle
forming the beam cross section. Then by geometry $r = a \sqrt(3)$.
\begin{align*}
  \text{rail} &=  1 \\
  \text{one-hop} &= \sqrt{\frac{1}{3}^2 + (\frac{a}{\sqrt{3}})^2 + (a/2)^2}\\
\text{one-hop}  &= \sqrt{\frac{1}{9} + \frac{7 a^2}{12}} \\
    \text{two-hops} &= \sqrt{\frac{2}{3}^2 + (\frac{a}{\sqrt{3}})^2 + (a/2)^2}  \\
\text{two-hops}    &= \sqrt{\frac{4}{9} + \frac{7 a^2}{12}}
\end{align*}

@@

We seek the value of $a$ that minimizes the maximum difference between these three lengths. We find
by numerical experimentation that this difference is minimized that the one-hop distance is equal to $1$,
which also has the pleasing property of giving us only two lengths. Then
\begin{align*}
   1  &=  \sqrt{\frac{1}{9} + \frac{7a^2}{12}} \\
     a  &=  \sqrt{\frac{8 \cdot 12}{9 \cdot 7}} \\
          a  &=  \sqrt{\frac{32}{21}}
\end{align*}

Therefore $ r = \sqrt{\frac{32}{63}} \approx 0.71267 $, producing length difference in the two-hop rail of
$\frac{2}{\sqrt{3}}$
and $1$ of $\approx 15.47\% $.

Note: Another interesting solution occurs at $a = \sqrt(5)/2$, which produces three length
classes of $11/12, 12/12, 13/12$.

\begin{conjecture}
  The equitetrabeam is also optimal under the metric of the total length of all members 
  for an equilateral tetrabream isomorphic to the Boerdijk--Coxeter helix.
\end{conjecture}

@@

We can then create a continuum of radius, length, rho to place into our general equation
to create continuum of helices between the BC and equitetrabeam. This continuum is 
minimax-length optimal at its endpoints, and may be optimal throughout.

The equitetrabeam has chirality.

@@

\begin{theorem}

  Using the six member classes defined by the non-primary colors, it is possible to smoothly twist and untwist a tetrahelix by
  using a linear combination of lengths.

\end{theorem}

@@

@@

The most important thing to do here is to start a github project that includes javascript
to copute this stuff; the second most important thing is to use Three.js to create a
nice visual representation of all of this without getting the physics engine involved.

Check formula to make sure lengths are what we previously created for the Equitetrabeam.

First, plug values into Equitetrabeam.

Check distances.

Also, update diagam.

@@

  Investigate Lambda outsidde the range [0,1].  Does -lambda reverse the chirality?
  Does lambda > 1 increase twist nicely?

  Create alternate coloring inside my system to render the tetrabot better.
  Add shadows better.

  Conjecture: There is a sense in which every tetrahelix has vertices
  evenly space along the axis of motion in order to have optimal lengths.

  Argument sketch: Follow current argument sketch, but compute lengths from formula
  for helices, using symmetry argument whereever possible.

  Create separate Github project having javascript code as function for computing all.

>>>>

\begin{abstract}
  The Boerdijk--Coxeter helix (BC helix, or tetrahelix) is a face-to-face stack of regular tetrahedra.
  Considering the edges of these tetrahedra, the resulting structure is attractive and inherently rigid,
  and therefore interesting to architects, mechanical engineers,
  and robotocists.
  A formula that matches the visually apparent helices forming the outer rails of the tetrahelix is derived which
  is convenient for designers.
  This formula defines the vertices of tetrahelices of varying radius, pitch, and curvature, with the BC helix
  as a special case. 
  An addtional special case is of that of $0$ curvature or rail angle, which generates a \emph{tetrabeam}.
  A particular choice of paramaters defines a novel object, the  minimax member length-difference tetrabeam,
  the \emph{equitetrabeam}.
  Linear interpolation of these parameters with the BC-helix parameters defines a continuum of close-to regular
  tetrahelices of designable curvature, pitch based on a single parameter.
  Utility and use for static and variable geometry truss/space frame design and robotics based on the are discussed.
\end{abstract}

@@

\begin{theorem}
  \label{tetrahelixoptimality}
  Any one tetrahedron in a minimax-optimal tetrahelix $ABCD$ of edge lengths $\overline{\rm AD} = 1, \overline{\rm AB} = \overline{\rm BC} = \overline{\rm CD} = u , \overline{\rm AC} = \overline{\rm BD} = v$, where $1 \leq u \leq v$.
  Any one tetrahedron in a tetrahelix has $1$ rail edge, $2$ mid-length edges connected to the rail and $2$ long edges connected to the rail.
  The edge opposite of the rail edge is a mid-length edge.
\end{theorem}

@@

Note that based on Theorems \ref{tetrahelixoptimality} and \ref{eventhirds}, we are justified in classifying edge lengths as \emph{rail},\emph{one-hop}, or
\emph{two-hops}. The one-hop edges are the edges between closest on the $z$-axis, and the two-hop edges are those that hop over a vertex.

@@

@@

Checking the important special case of the BC helix, we find that this equation
indeed holds true (treating $d$ in this equation as $3 h_{bc}$ as defined by
Gray and Coxeter), where they are using it for the axial height from one node to
the next of a different color, but we use it to mean distance for the same color.

@@

By Theoerm \ref{eventhirds}, we can compute the (at most) three edge-lengths of an optimal
tetrahelix by (where $dist$ is the cartesian distance function):
\begin{align*}
  \text{rail} &= dist(H_{general}(n,c,\rho,h_{\rho}),H_{general}(n+1,c,\rho,d_{\rho})) = 1 \\
  \text{one-hop} &= dist(H_{general}(n,c,\rho,h_{\rho}),H_{general}(n,c+1,\rho,d_{\rho}))  \\
  \text{two-hops} &= dist(H_{general}(n,c,\rho,h_{\rho}),H_{general}(n,c+2,\rho,d_{\rho}))  \\  
\end{align*}
Where are invarinat for all $n$ and $c$.

@@

 \begin{figure}[H]
     \centering
     \includegraphics[width=0.9\textwidth]{figures/EquilateralDiagram.png}
     \caption{Coloring of an (untwisted) Tetrabeam}
 \end{figure}

\begin{theorem}
The equitetrabeam with minimal maximal edge difference is produced by $H_{general}$ when $ r = \sqrt{\frac{8}{27}} $.
  \end{theorem}

When $\rho = 0$, our formulae for distances are particular simple. Let $a$ be the length of the equilateral triangle
forming the beam cross section. Then by geometry $r = a \sqrt{3}$.
\begin{align*}
  \text{rail} &=  1 \\
  \text{one-hop} &= \sqrt{\frac{1}{3}^2 + (\frac{3a}{2\sqrt{3}})^2 + (a/2)^2}\\
\text{one-hop}  &= \sqrt{\frac{1}{9} + a^2} \\
    \text{two-hops} &= \sqrt{\frac{2}{3}^2 + (\frac{3a}{2\sqrt{3}})^2 + (a/2)^2}  \\
\text{two-hops}    &= \sqrt{\frac{4}{9} + a^2}
\end{align*}
Computing the derivative of two-hops - one-hop:
\begin{align*}
 \frac{\partial \text{two-hops} - \text{one-hop}}{a} &= \frac{\partial \sqrt{4/9 + a^2} - \sqrt{1/9 + a^2}}{\partial a} \\
  \frac{\partial \text{two-hops} - \text{one-hop}}{a} &= \frac{a}{\sqrt{a^2 + 4/9}} - \frac{a}{\sqrt{ a^2 + 1/9}} 
\end{align*}
which is negative for all $a> 0$. Thererfore two-hop - one-hop
is minimized when we maximize $a$, up until one-hop = rail.

@@

We seek the value of $a$ that minimizes the maximum difference between these three lengths. We find
by numerical experimentation that this difference is minimized that the one-hop distance is equal to $1$,
which also has the pleasing property of giving us only two lengths. Then
\begin{align*}
   1  &=  \sqrt{\frac{1}{9} + a^2} \\
     a  &= \sqrt{\frac{8}{9}} 
\end{align*}

Therefore $ r = \sqrt{\frac{8}{27}} \approx 0.5443 $, producing length difference in the two-hop rail of
$\frac{2}{\sqrt{3}}$
and $1$ of $\approx 15.47\% $.

Note: Another interesting solution is derived by setting (one-hop + two-hop)/2 = 1,  occurs at $a = \sqrt{35/3}/4$,
which produces three length classes of $11/12, 12/12, 13/12$.

\begin{conjecture}
  The equitetrabeam is also optimal under the metric of the total length of all members 
  for an equilateral tetrabeam isomorphic to the Boerdijk--Coxeter helix.
\end{conjecture}

@@

We seek to unify this with degenerate helix formula for the equitetrabeam:
\[
H_{etb}(n,c) =
\left [
  \begin{tabular}{c}
   $ r_{etb}  \cos( 0 \cdot \kappa  + c 2 \pi /3) $\\
   $ r_{etb}  \sin( 0 \cdot \kappa  + c 2 \pi /3) $\\
   $ \kappa d_{etb} $
  \end{tabular}
\right ],
\text{where:}
  \begin{tabular}{c}
 $\kappa = n + c/3$ \\
    $\rho_{bc} = (3 \theta - 2 \pi)$ \\
   $ \theta = \arccos(-2/3) $
  \end{tabular}      
\]
where $ r_{etb} = \sqrt{\frac{8}{27}}$, $d_{etb} = 1$,

\begin{align*}
r_{\lambda}  &=  \lambda (\frac{3 \sqrt{3}}{10}  - \sqrt{\frac{8}{27}}) + \sqrt{\frac{8}{27}}  \\
d_{\lambda} &=   \lambda (3 \sqrt{10} - 1) + 1 \\
\phi_{\lambda} &=  \lambda \rho_{bc}  + 0
\end{align*}
to create a formula that generates a continuum of tetrahedral structures:

\[
H_{continuum}(n,c,\lambda) =
\left [
  \begin{tabular}{c}
   $ r_{\lambda} \cos(\phi_{\lambda} \kappa + c 2 \pi /3) $\\
   $ r_{\lambda}  \sin(\phi_{\lambda} \kappa + c 2 \pi /3) $\\
   $ d_{\lambda} \kappa $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
    $\kappa = n + c/3$ \\
    $ \phi_{\lambda} =  \lambda \rho_{bc}  + 0 $\\
    $\rho_{bc} = (3 \theta - 2 \pi)$ \\
   $ \theta = \arccos(-2/3) $
  \end{tabular}      
\]
A value of $\lambda = 0$ generates the equitetrabeam, and $\lambda = 1$ generates the Boerdijk--Coxeter helix, and every
value $\lambda \in [-1,1]$ generates an attractive structure which if physically realized is an inherently rigid structure
with member lengths of no greater disparity than $16\%$.
In the sense of structural engineering, it would be a relatively strong space frame.

 \begin{figure}[H]
     \centering
     \includegraphics[width=0.9\textwidth]{figures/Continuum.png}
     \caption{A Continuum of Tetrahelices}
 \end{figure}

Negative values of $\lambda$ generate clockwise tetrahelices; positive generate counter-clockwise helices.

Furthermore, this formula allows one to design the pitch (in edge length units) as a function of $\lambda$ of the helix (along one rail)
where $\lambda \neq 0$:
\[
p(\lambda) = 2 \pi  \cdot \frac{(\frac{8}{27} -1) \lambda +1)}{ \lambda  \rho_{bc} }
\]

We can then create a continuum of radius, length, rho to place into our general equation
to create continuum of helices between the BC and equitetrabeam. This continuum is 
minimax-length optimal at its endpoints, and may be optimal throughout. This untwisting is accomplished
only by changing the length of the two-hop member.

@@

\begin{theorem}
  By changing only the length of the members that change rails and make two nodes, you can untwist a tetrobot
  form the Boerdijk-Coxeter configuration to the equitetrabeam which rests flat on the plane.
\end{theorem}

@@

See if the continuum holds the short edge lengths at one perpetually.  Yes!!!

Create working toolkit for desin/exploration. Improve code that way, use white background, floor.

Try to compute optimal radius within our context. This would let us assert that we have
an ``optimum'' continuum.  Not worth much to an engineer, but valuable.

@@

@@
